---
output:
  html_document: default
  pdf_document: default
---


```{r, echo = FALSE, results='hide', message=FALSE}
library(knitr)
library(knitLatex)
Sys.setlocale("LC_TIME", "C")
opts_chunk$set(echo = TRUE)
```


# **Setting Working Directory And Loading the Data**
***



```{r}
if(!file.exists("./data")){dir.create("./data")}
setwd("C:/Users/Heosh/Desktop/data/JHU_reproducibleResearch/data")
activity <- read.csv("activity.csv")
str(activity)
head(activity)
```
```{r, echo = FALSE, results='hide', message=FALSE}
library(dplyr)
library(tidyr)
```



# **What is mean total number of steps taken per day?**
***



#### Q1) Calculate the total number of steps taken per day

```{r}
tapply(activity$steps, activity$date, sum)
```


#### Q2) Make a histogram of the total number of steps taken each day

```{r}
step1 <- tapply(activity$steps, activity$date, sum)

hist(step1, breaks = 20, xlab = "Total Number of Steps Each Day", main = "Total Number of Steps Taken Each Day")
```

#### Q3) Calculate and report the mean and median of the total number of steps taken per day

```{r}
tapply(activity$steps, activity$date, mean)
mean(activity$steps, na.rm = T)

tapply(activity$steps, activity$date, median)
median(activity$steps, na.rm = T)
```

# **What is the average daily activity pattern?**
***


#### Q1) Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days

```{r}
activity$date <- as.Date(activity$date, "%Y-%m-%d")

date1 <- tapply(activity$steps, activity$date, mean, na.rm = T)

plot(date1, type = "l", xlab = "Interval", ylab = "Steps", main = "Interval and Steps")
```

##### Q2) Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
int1 <- tapply(activity$steps, activity$interval, mean, na.rm = T)

int1[which.max(int1)]

```

# **Imputing missing values!**
***


#### Q1) Calculate and report the total number of missing values in the dataset

```{r}
sum(as.numeric(is.na(activity$steps)))
sum(as.numeric(!is.na(activity$steps)))
```

#### Q2), Q3) Devise a strategy for filling in all of the missing values in the dataset. & Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r, echo = FALSE, results='hide', message=FALSE}
library(Hmisc)
```

```{r}
activity2 <- activity

activity2$steps <- impute(activity2$steps, mean)
```

#### Q4) Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.


```{r}
step2 <- tapply(activity2$steps, activity2$date, sum)
hist(step2, breaks = 20, xlab = "Total Number of Steps Each Day", main = "Total Number of Steps Taken Each Day")
```


```{r}
tapply(activity2$steps, activity2$date, mean)
mean(activity2$steps)

tapply(activity2$steps, activity2$date, median)
median(activity2$steps)
```


# **Are there differences in activity patterns between weekdays and weekends?**
***

#### Q1) Create a new factor variable in the dataset with two levels – “weekday” and “weekend”

```{r}
activity3 <- mutate(activity2, weekday = weekdays(activity2$date))
activity4  <- mutate(activity3, day = ifelse(activity3$weekday == "Saturday" | activity3$weekday == "Sunday", "Weekend","Weekday"))

activity4$day <- as.factor(activity4$day)
```
str(activity4)


#### Q2) Make a panel plot containing a time series plot of the 5-minute interval and the average number of steps taken, averaged across all weekday days or weekend days.

```{r, echo = FALSE, message = FALSE, results = "hide"}
library(ggplot2)
```

```{r}

actPlot <- summarise(group_by(activity4, interval, day), steps = mean(steps))
actPlot

ggplot(actPlot, aes(x = interval, y = steps))+ geom_line() + facet_grid(.~ day) + labs(x ="Intervals", y = "Average Steps")


```




